---
output: github_document
always_allow_html: true
---

# njmunicipalities-api

This is the source code for the NJ Municipalities API deployed at `api.tor-gu.com`.

It contains the names and GEOID/FIPS codes for every municipality and county in NJ, from 2000 to 2022, and records the changes from year to year.

## Try it out

### OpenAPI/Swagger
There is an OpenAPI/swagger document deployed [here](https://tor-gu.com/apis/njmunicipalities.html)
that you can play with.


### Sample code
Each endpoint returns paginated results, with a `"next"` url in the `"links"` section 
indicating that there are more results to be retrieved.

This is illustrated in the python code below, which loads the table of municipalities for 2022 
into a [pandas DataFrame](https://pandas.pydata.org/docs/reference/api/pandas.DataFrame.html).
```{python}
import json
import requests
import pandas as pd


def load_as_df(url):
	"""Retrieve entire dataset at url and return as a dataframe"""
    # Get the first page
	resp = requests.get(url)
	resp.raise_for_status()
	resp_json = json.loads(resp.content.decode('utf-8'))
	df = pd.DataFrame(resp_json["data"])
    
    # Keep retrieving pages while there is a 'next' link
	while "next" in resp_json["links"]:
		resp = requests.get(resp_json["links"]["next"])
		resp.raise_for_status()
		resp_json = json.loads(resp.content.decode('utf-8'))
		df = pd.concat([df, pd.DataFrame(resp_json["data"])], ignore_index = True)
	return df


# Load municipality tables for 2022
municipalities_2022 = load_as_df("https://api.tor-gu.com/nj/municipalities/2022")
```
```{r echo=FALSE}
library(reticulate)
knitr::kable(head(py$municipalities_2022))
```

The `municipality_xrefs` endpoint works similarly. As an example, let's also retrieve the 
2000 municipality table along with the 2022/2000 cross-reference table, and take a look
at the changes in NJ municipalties between 2000 and 2022.
```{python comment="", strip.white=TRUE}
municipalities_2000 = load_as_df("https://api.tor-gu.com/nj/municipalities/2000")
xrefs = load_as_df("https://api.tor-gu.com/nj/municipality_xrefs/2022/2000")
# Select columns for a merged table
xrefs = xrefs.rename(
	columns={"GEOID_ref": "GEOID_2022", "GEOID": "GEOID_2000"}
	)[["GEOID_2000", "GEOID_2022"]]
municipalities_2000 = municipalities_2000.rename(
	columns={"GEOID": "GEOID_2000", "municipality": "municipality_2000"}
	)[["GEOID_2000", "county", "municipality_2000"]]
municipalities_2022 = municipalities_2022.rename(
	columns={"GEOID": "GEOID_2022", "municipality": "municipality_2022"}
	)[["GEOID_2022","municipality_2022"]]

# Create the merged table
merged = xrefs.merge(
	municipalities_2000, how="left", on="GEOID_2000").merge(
	municipalities_2022, how="left", on="GEOID_2022").fillna("NA")

# Select all changes in municipalities between 2000 and 20222
changes = merged[(merged["municipality_2000"] != merged["municipality_2022"]) | 
	(merged["GEOID_2000"] != merged["GEOID_2022"])]
```
```{r echo=FALSE}
knitr::kable(py$changes)
```

## Deploying
This API is implemented as an [AWS SAM](https://aws.amazon.com/serverless/sam/) application. It depends on a lambda layer
[layer_torguapi](https://github.com/tor-gu/layer_torguapi), which needs to be deployed separately.  You also 
need to have CSV dumps of the tables in [njmunicipalities](https://github.com/tor-gu/njmunicipalities) available
in an S3 bucket.

### Direct deploy
The application can be deployed directly using the SAM CLI:
```
sam build
sam deploy --stack-name MyStackName
           --template template.yaml
           --parameter-overrides                                                            \
                TorguapiLayerArn=arn:aws:lambda:us-east-1:123456789012:layer:layer_torguapi \
                TorguapiLayerVersion=1                                                      \
                ApiRoot=https://api.example.com 
```
Replace the `TorguapiLayerXXX` params with the values derived from the `layer_torguapi` deployment.

The `ApiRoot` value is used only for generating the values in `"links"` in the JSON reply. On 
`api.tor-gu.com`, the relationship between the publicly facing API URL and the API Gateway 
endpoints is configured in [torgu-api-cloudfront](https://github.com/tor-gu/torgu-api-cloudfront).

### Using a pipeline
The package also include the code pipeline that I use to deploy from github.

In addition to the requirements above, you will need to have execution roles for the 
pipeline (`PipelineExecutionRole`) and for the deployment itself 
(`CloudFormationExecutionRole`).

```
aws cloudformation deploy --stack-name MyPiplineStackName \
	--template-body file://codepipeline.yaml 					\
	--capabilities CAPABILITY_IAM								\
	--parameters \
        ParameterKey=ArtifactBucket,ParameterValue=MyArtifactBucket \
		ParameterKey=PipelineExecutionRole,ParameterValue=arn:aws:iam::123456789012:role/MyPipelineExecutionRole \
		ParameterKey=CloudFormationExecutionRole,ParameterValue=arn:aws:iam::123456789012:role/MyCloudFormationExecutionRole \
		ParameterKey=Region,ParameterValue=us-east-1 \
        ParameterKey=DeployStackName,ParameterValue=MyStackName \
        ParameterKey=TorguapiLayerArn,ParameterValue=arn:aws:lambda:us-east-1:123456789012:layer:layer_torguapi \
        ParameterKey=TorguapiLayerVersion,ParameterValue=1 \
        ParameterKey=ApiRoot,ParameterValue=https://api.example.com
```
